---
title: "Exploratoring environmental data"
author: "RJ"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, encoding = "UTF_8")


library(ggplot2)
library(lubridate)
library(sf)
library(shiny)
library(htmlwidgets)
library(gapminder)
library(plotly)
library(dplyr)
library(tidyr)
library(maptools)
library(data.table)
library(Imap)
library(httr)
library(ncdf4) # to get what's needed from ncdf files

# set to fractional seconds
op <- options(digits.secs=3)

# Colony
# from Alex Corbeau who did fieldwork there. 46°21'31.1"S 51°42'25.1"E 
colony <-c(51.706972, -46.358639)

# let's do this around infrasound and barometric pressure
path_to_env_files <- "./Databse Ollie/"
ID_folders <- dir(path_to_env_files, pattern = "ID")

# let's do this for the 1st bird
x <- 7 # first ID
path_to_trips <- paste0(path_to_env_files, ID_folders[x], "/")
ID_trips_folders <- dir(path_to_trips)

x_trip <- 2 # second trip for the ID
path_to_ncdf <- paste0(path_to_trips, ID_trips_folders[x_trip], "/netCDF/")

```

## Pressure

```{r Pressure , echo=FALSE, include = F}
# open Barometric pressure file
baro_name <- dir(path_to_ncdf, pattern = "Barometric")
nc_file <- nc_open(paste0(path_to_ncdf, baro_name))
print(nc_file) # getting all the characteristics of the file
Pressure_matrix <- t(ncvar_get(nc_file, "Full_matrix"))
nc_close(nc_file)
Pressure_df <- as.data.frame(Pressure_matrix)
colnames(Pressure_df) <- c("Time","Lon","Lat","Pressure_hPa")
str(Pressure_df)
Pressure_df$Time <- as.POSIXct(Pressure_df$Time, tz="GMT", origin = "1970-01-01 00:00:00") + hours(5) # adding time difference with Crozet
head(Pressure_df$Time) 
```

### Some context for each track's pressure data 

* Source: Ollie's notebooks and email exchanges between Jelle, Ollie and me

* The data are not calibrated yet.
* Since there should be more or less 5-min segments of data every hour, the raw data at 10 Hz are first divided into 5 minute segments (or 15, but it will only contain 5)
* Data cleaning is done on these segments. So, each segments:
  * Assuming that barometric pressure distributes normally --Jelle said it's a safe assumption --, outliers are filtered out by first obtaining z-scores
  and keeping the observations whose absolute z-scores are below 3. 
  * Then, if consecutive observations show a difference greater than 2hPa (meaning a ~20m difference),
  the following value is replaced by the preceding value. 
* Resampling is done at 1 Hz. It is basically done by fitting sinusoids to the data (Fast Fourier Transform) and then taking value points. We're using the obspy.resample function in python which is a wrapper of scipy.signal.resample
* Everything is saved into a netCDF file


### With GPS

```{r gpsPressurePlot , echo=FALSE}
###create an interactive figure of lat long using ggplotly with the colony point
data <- Pressure_df
lon<-c(data$Lon,51.706972)
lat<-c(data$Lat,-46.358639)
p0 <- ggplot(data, aes(Lon, Lat)) +
  geom_point(aes(color = Pressure_hPa),size=0.5)+
  geom_text(x=colony[1], y=colony[2], label='Colony')+
  geom_point(x=colony[1], y=colony[2], color="red")+
  scale_colour_viridis_c(option = "magma", direction = -1) +
  xlim(min(lon), max(lon))+
  ylim(min(lat), max(lat)) +
  theme_bw()
ggplotly(p0,
         dynamicTicks = TRUE, session="knitr")
```

### Time series
```{r PressureTS , echo=FALSE}
###create an interactive figure of lat long using ggplotly with the colony point
pPresTS <- data %>%
  ggplot(aes(x =Time, y = Pressure_hPa)) +
  geom_point(aes(color = Pressure_hPa),size=0.5)+
  scale_colour_viridis_c(option = "magma", direction = -1) +
  theme_minimal()
ggplotly(pPresTS,
         dynamicTicks = TRUE, session="knitr")

```

## Infrasound

### Some context for each track's infrasound data

* Source: Ollie's notebooks, email exchanges between Jelle, Ollie and me, some reading and youtube videos

* The data are calibrated.
* Since there should be more or less 5-min segments of data every hour (could be 1), the raw data at 10 Hz are first divided into 5 minute segments (or 15, but it will only contain 5)
* Data cleaning is done on these segments. So, each segments:
  * Assuming that absolute pressure distributes normally, outliers are filtered out by first obtaining z-scores
  and keeping the observations whose absolute z-scores are below 3. 
  * Then, if consecutive observations show a difference greater than 10 (I don't know why 10),
  the following value is replaced by the preceding value. 
* No resampling
* For each segment, the power spectral density is estimated using Welch's method;
  it basically divides data into overlapping segments, computing a modified periodogram for
  each segment and averaging the periodograms
* The PSD data are transformed to obtain SPL as $10 \times \log(PSD/(2*10^{-5})*2))$
* The SPL data are saved into a netCDF file

```{r Infrasound , echo=FALSE, include = F}
# open Infrasound file
sound_name <- dir(path_to_ncdf, pattern = "Sound")

nc_file <- nc_open(paste0(path_to_ncdf, sound_name))
print(nc_file) # getting all the characteristics of the file

# getting the names of the columns
names_var <- names(nc_file$var)
names_var <- names_var[1:(length(names_var)-1)]
SPL_Matrix <- ncvar_get(nc_file, "Full_matrix")
nc_close(nc_file)
SPL_df <- as.data.frame(SPL_Matrix)
colnames(SPL_df) <- c("Time",names_var[2:length(names_var)])
str(SPL_df)
Freq <- SPL_df[1,4:dim(SPL_df)[2]] # These are the frequencies for SPL
SPL_df <- SPL_df[-1,]
SPL_df$Time <- as.POSIXct(SPL_df$Time, tz="GMT", origin = "1970-01-01 00:00:00") + hours(5)
# We may not be interested in each frequency for now
# 

# We'll make two categories of SPL: 0.07-1 and 1-5
SPL_low <- apply(SPL_df[,which(Freq < 1 & Freq > 0.07)+3],1,mean)
SPL_high <- apply(SPL_df[,which(Freq > 1)+3],1,mean)

SPL_cat <- cbind.data.frame(SPL_df[,1:3], SPL_low, SPL_high)
```


### GPS - Sound Pressure Level from Low Infrasound (0.07-1Hz)
```{r gpsSPLlowPlot , echo=FALSE}
###create an interactive figure of lat long using ggplotly with the colony point
data <- SPL_cat
lon<-c(data$Lon,51.706972)
lat<-c(data$Lat,-46.358639)
pGPSlow <- ggplot(data, aes(Lon, Lat)) +
  geom_point(aes(color = SPL_low),size=0.5)+
  geom_text(x=colony[1], y=colony[2], label='Colony')+
  geom_point(x=colony[1], y=colony[2], color="red")+
  scale_colour_viridis_c(option = "magma", direction = -1) +
  xlim(min(lon), max(lon))+
  ylim(min(lat), max(lat)) +
  theme_bw()
ggplotly(pGPSlow,
         dynamicTicks = TRUE, session="knitr")
```

### Time series of Sound Pressure Level from Low Infrasound (0.07-1Hz)
```{r SPLlowTS , echo=FALSE}
###create an interactive figure of lat long using ggplotly with the colony point
pSPLlowTS <- data %>%
  ggplot(aes(x =Time, y = SPL_low)) +
  geom_point(aes(color = SPL_low),size=0.5)+
  scale_colour_viridis_c(option = "magma", direction = -1) +
  theme_minimal()
ggplotly(pSPLlowTS,
         dynamicTicks = TRUE, session="knitr")

```

### GPS - Sound Pressure Level from High Infrasound (1-5Hz)
```{r gpsSPLhighPlot , echo=FALSE}
###create an interactive figure of lat long using ggplotly with the colony point
data <- SPL_cat
lon<-c(data$Lon,51.706972)
lat<-c(data$Lat,-46.358639)
pGPShigh <- ggplot(data, aes(Lon, Lat)) +
  geom_point(aes(color = SPL_high),size=0.5)+
  geom_text(x=colony[1], y=colony[2], label='Colony')+
  geom_point(x=colony[1], y=colony[2], color="red")+
  scale_colour_viridis_c(option = "magma", direction = -1) +
  xlim(min(lon), max(lon))+
  ylim(min(lat), max(lat)) +
  theme_bw()
ggplotly(pGPShigh,
         dynamicTicks = TRUE, session="knitr")
```

### Time series of Sound Pressure Level from High Infrasound (1-5Hz)
```{r SPLhighTS , echo=FALSE}
###create an interactive figure of lat long using ggplotly with the colony point
pSPLhighTS <- data %>%
  ggplot(aes(x =Time, y = SPL_high)) +
  geom_point(aes(color = SPL_high),size=0.5)+
  scale_colour_viridis_c(option = "magma", direction = -1) +
  theme_minimal()
ggplotly(pSPLhighTS,
         dynamicTicks = TRUE, session="knitr")

```
